---
title: "Prosper Loan Analysis"
author: "James Malcolm"
output:
  html_document:
    df_print: paged
---

This Markdown and repo is dedicated to exploring Prosper Loan data to answer
the following questions:
* Who uses Prosper/Borrowers



```{r}
# Load Packages needed
library(ggplot2)

```

Now we read the data set in, and check out a few statistics.
```{r}
loans <- read.csv('./data/prosperLoanData.csv')

# View shape of data
dim(loans)
```

```{r}
str(loans)
```

I note, that Terms is not represented as a factor. However we know from Prosper
that Terms of Loans are only 12, 36, or 60 months. So we change this column to
be a Factor data type.

```{r}
# Set Term data type to be a factor
loans$Term <- factor(loans$Term,
                     levels = c(12,36,60),
                     ordered = TRUE)
```

An initial plot that I want to make is to simply view the distribution of
loan amounts that have been issued. I do this using a Histogram.

```{r}
# Plot histogram of Loan amounts
ggplot(aes(x = LoanOriginalAmount), data = loans) + 
  geom_histogram(binwidth = 1000) +
  xlab('Original Loan Amount (USD)') +
  ggtitle('Histogram of original loan amount')
```

View this graph is interesting, as we can see that there are spikes in loan
amounts at round numbers, eg $10,000, $15,000, $20,000, etc. This reflects that
humans tend to think in round numbers as opposed to requestings loans for
specific number, eg. $7824.

We can also see that while most the majority of loans are $25,000 or less, there
are a number of loans above this figure.

A 
method to assess wehether these are valid loans or not would be see the reason
as to why they're getting the loan! Since however, we don't have have the
reason, we're going to continue doing more EDA on it.

```{r}
min(loans$LoanOriginalAmount)
max(loans$LoanOriginalAmount)
```

Further research on Prosper's Help Centre finds that Prosper only allows loans
ranging from $2,000 to $35,000. Given we have no datapoints beyond $35,000, I'm 
confident enough that these potential outliers abover $25,000 are legitamate 
loans - so I won't remove or otherwise transform them.

An assumption that I want to visualise is that the larger the Loan amount, the
larger the Term would be. To view this, I create the following boxplote to
explore the relationship of loan amounts across the three different terms 
(12 months, 36 months and 60 months).

```{r}
ggplot(aes(x = Term, y = LoanOriginalAmount), data=loans) +
  geom_boxplot() +
  xlab('Term (Months)') +
  ylab('Original Loan Amount (USD)') +
  ggtitle('Distribution of Loan Amounts per Term')

```



